---
chapter: "none yet"
title: "Title for cow-type-kayak"
author: "Danny Kaplan"
output: learnr::tutorial
tutorial:
  id: "cow-type-kayak"
  version: 0.1
runtime: shiny_prerendered
date: 2020-06-04 
tags: [half life, exponentials, third]
id: cow-type-kayak
---

```{r cow-type-kayak-setup, include = FALSE}
SDSdata::sds_setup()
library(etude)
library(learnr)
library(gradethis)

library(ggformula)
library(mosaic)
library(mosaicCalc)
library(mosaicUSAFA)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = QandA::save_check_inputs)
```


`r etude::exercise_title()` The Wikipedia entry on "Common Misconceptions" used to contain this item:

> *Some cooks believe that food items cooked with wine or liquor will be non-alcoholic, because alcohol's low boiling point causes it to evaporate quickly when heated. However, a study found that some of the alcohol remains: 25% after 1 hour of baking or simmering, and 10% after 2 hours.*


```{r cow-type-kayak-C1}
etude::essay_response("a) State briefly ways in which the given information is or is not consistent with an approximately exponential process.")
```

````{etude}
The amount of alcohol remaining in the food is decreasing toward zero.  The measurement of 25% after 1 hour suggests a half-life of 30 minutes, but the decrease from 25% to 10% in the second hour indicates that the half-life is somewhat different, a bit more than 30 minutes.
````



  
```{r cow-type-kayak-Qinline1, echo=FALSE}
explain <- "These are not as required within $\\pm 2.5$ percentage points of the numbers given in Wikipedia."
etude::choose_one(
  prompt = "b) The stated measurements appear rounded to about 5 percentage points --- 25%, 10%.  Give a small alteration to the stated measurements (to within $\\pm 2.5$ percentage points) that would be exactly consistent with an exponential process.?",
  c("Decrease to 23.6% after one hour and 5.6% after two." = explain,
    "+Decrease to 27.5% after one hour and 7.6% after two.+" = "",
    "Decrease to 40% after one hour and 24% after two." = explain, 
    "Decrease to 20% after one hour and 10% after two." = "The decrease in the first hour is by 80%, corresponding to a half life of about 26 minutes. But in the second hour the half life is 60 minutes. These are inconsistent."),
  random_answer_order = FALSE,
  inline=FALSE
)
```

```{etude}
Suppose that the decrease in the first hour were to 27.5% which is 2.5 percentage points different from 25%.  Decreasing at exactly this rate for the next hour would result in $0.275^2$ remaining after two hours, that is, $0.0756 = 7.56$%, which is also within 2.5 percentage points of the stated value.
```

