---
chapter: "none yet"
title: "Title for tiger-have-fork"
author: "Danny Kaplan"
difficulty: "XX"
output: learnr::tutorial
tutorial:
  id: "tiger-have-fork"
  version: 0.1
runtime: shiny_prerendered
date: 2020-07-19 
tags: [mortality, logs, third]
id: tiger-have-fork
---

```{r tiger-have-fork-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(math141Zexercises)
library(ggformula)
library(ggplot2)
library(mosaicCalc)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` The data frame `SSA_2007` comes from the [US Social Security Administration(https://www.ssa.gov/oact/STATS/table4c6.html) and contains mortality in the US as a function of age and sex. ("Mortality"  refers to the probability of dying in the next year.)

```{r tiger-have-fork-sandbox1, exercise=TRUE, exercise.cap="Sandbox",  exercise.eval=FALSE, exercise.lines=2, exercise.completion=FALSE, exercise.startover=TRUE, exercise.diagnostics=TRUE}

```

The next chunk would make a plot of the relationship, but there is a *slight* mistake in the way the command is written. Use the sandbox to explore the `SSA_2007` data frame using functions such as `head()` and `names()` to gain some insight into what the problem is with the plotting command.  


```{r tiger-have-fork-E1, exercise=TRUE, exercise.cap="Sandbox", exercise.eval=FALSE, exercise.lines=2, exercise.completion=FALSE, exercise.startover=TRUE, exercise.diagnostics=TRUE}
gf_point(mortality ~ age, color = ~ sex, data = SSA_2007)
```

```{r tiger-have-fork-QA1, echo=FALSE}
etude::choose_one(
  prompt = "A) What was the mistake in the plotting command in the above code box?",
  choices = list(
    "+Variable names didn't match the ones in the data.+" = "",
    "The *tilde* in the argument `color = ~ sex`" = "The `color =` argument is right. The value being used, `~ sex`, is a one-sided formula and is used for things like color, shape, transparency, ....",
    "The data frame name is spelled wrong." = "No.",
    "There is no function `gf_point()`." = "No. `gf_point()` is one of the more commonly used plotting functions"
  )
)
```

```{r tiger-have-fork-Essay1, echo=FALSE}
etude::essay_response(
  prompt = "B) What's the obvious message of the above plot?"
)
```

Now you are going to use semi-log and log-log scales to look at the mortality data again. To do this, you will use the `gf_refine()` function. As the name suggests, this will look like:

```r
gf_point( put plotting arguments here ) %>% 
  gf_refine()
```

To make the y-axis show up as a log scale, use as an argument to `gf_refine()` the value produced by the `scale_y_log10()` function. Similarly, use (as another argument to `gf_refine()`) the value produced by  `scale_x_log10()` to put the x-axis on a log scale.

Modify the plotting command to make a semi-log plot of mortality versus age. Interpret the plot to answer the following questions. Note that labels such as those along the vertical axis are often called "decade labels."

```{r tiger-have-fork-QA2, echo=FALSE}
message = "C) Hint: How much is the change between successive labels on the y axis?"
etude::choose_one(
  prompt = "The level of mortality in year 0 of life is how much greater than in year 1 and after?",
  choices = list(
    "About twice as large." = message,
    "About five times as large" = message,
    "+About 10 times as large+" = "",
    "About 100 times as large" = message
  )
)
```

```{r tiger-have-fork-QA3, echo=FALSE}
message = "Hint: Due to the nature of logs, a difference of half a decade corresponds to a change of $\\sqrt{10}$."
etude::choose_one(
  prompt = "D) Near age 20, the mortality of males is how much compared to females?",
  choices = list(
    "Less than twice as large." = message,
    "+A bit more than three times as large+" = "",
    "About 8 times as large" = message,
    "About 12 times as large" = message
  )
)
```

```{r tiger-have-fork-QA4, echo=FALSE}
etude::choose_one(
  prompt = "E) Between the ages of about 40 and 80, how does mortality change with age?",
  choices = list(
    "It stays about the same." = "But the curve is sloping up, isn't it?",
    "It increases as a straight line." = "It would be fair to say this about the logarithm of mortality. But a straight line in log mortality means that mortality itself is increasing exponentially.",
    "+It increases exponentially.+" = "",
    "It increases, then decreases, then increases again." = "Interesting that you would say this when the function in clearly monotonically increasing above age 30."
  )
)
```

Remake the plot of mortality vs age once again, but this time put it on log-log axes. The sign of a power-law relationship is that it shows up as a straight line on log-log axes. 

```{r tiger-have-fork-QA5, echo=FALSE}
etude::choose_one(
  prompt = "F) Between the ages of about 40 and 80 is the increase in mortality better modeled by an exponential or a power-law process?",
  choices = list(
    "Power-law" = "But it's hard to find any straight line on the log-log  plot.",
    "+Exponential+" = "Right. The graph is much closer to a straight line on semi-log scales than on log-log scales.",
    "No reason to prefer one or the other." = "One is much closer to a straight line than the other."
  )
)
```
