---
chapter: "none yet"
author: "Daniel Kaplan"
date: 2019-09-17 
version: 0.1
tags: [exercise, linear, exponential, linear combinations]
id: horse-spend-futon 
---

(ref:horse-spend-futon) Exercise in file horse-spend-futon

```{r horse-spend-futon-default, include = FALSE}
SDSdata::sds_setup()
```

TITLE GOES HERE: Consider world swimming records in the 100m freestyle race. 

```{r horse-spend-futon-1, echo = FALSE}
gf_point(time ~ year | sex,  data = mosaicData::SwimRecords)
```

The improvement in record time (that is, the decline) might be modeled linearly or exponentially. Note these aspects of constant, linear and exponential functions:

Constant
: Never changes

Linear
: If the slope is nonzero, the linear function must eventually cross zero.

Exponential
: Exponential decline must eventually reach zero (but doesn't cross).

Consider some basic facts about the possibilities for world record swim times:

i. Will they reach zero?
#. If so, will they cross zero?

1. How can you put together a linear combination of a constant and an exponential function in order to satify  your answers to i and ii.
2. What would be the parameters? Use `fitModel()` and do the calculations for males and females separately.
3. Based on your parameters, will the world record for women eventually reach or surpass that of men?


<!--begin-answer-->
```{r}
fitModel(time ~ A + B * exp(- (year-1950) / tau), data = mosaicData::SwimRecords %>% filter(sex == 'F'),
         start  = c(A  = 50,  B = 30,  tau=20) )
```
<!--end-answer-->

