---
chapter: "Quantities and their units"
author: "Daniel Kaplan"
date: 2019-09-18 
version: 0.1
tags: [notes, units, dimension]
id: pine-lose-clock 
---

(ref:pine-lose-clock) Exercise in file pine-lose-clock

```{r pine-lose-clock-default, include = FALSE}
SDSdata::sds_setup()
```

TITLE GOES HERE: Quantities and numbers

## What's a Unit?

Opening activity: Group the following in a natural way: foot, hour, mile, radian, dollar, liter, yen, second, day, euro, gallon, acre, hectare, cubic inch, lightyear, radian, teaspoon.

Comment: The groups correspond to different dimensions. The members of a single group are different units of measurement for the same dimension.

## What's a Dimension?


### Fundamental physical dimensions

Time, Length, Mass. T L M

### Common compound dimensions

-   area = L^{2}
-   volume = L^{3}
-   mass density = mass per volume = M L^{-3}
-   velocity = L T^{-1}
-   acceleration = velocity per T = L T^{-2}
-   force = M x acceleration = M L T^{-2}
-   energy = force x distance = M L^2 T^{-2}
-   power = energy per T = M L^2 T^{-2}
-   pressure = force per area = M L^{-1} T^{-2}
-   mass flux = mass per second per area = M L^{-2} T^{-1}

### Other common important dimensions

-   money (be it in units of dollars, euros, renmimbi, pounds, yen, rupee, rand, etc.)
-   temperature

### Units in economics

Economists write Y = f(K, L), production is a function of capital and labor. Production Y is taken to be a flow variable, units produced per unit time. Labor is usually also a flow, something like man-hours of labor per unit time. There are various conventions for K. It is simplest for students to take K to be in dollars, but you can amortize the capital or you can think of renting it over its lifetime, in which case K can also be thought of as a flow variable, dollars per unit time.

### Common, simple dimensionless quantities

-   angle
-   count
-   interest rate
-   percentage change of anything

Arithmetic with units
-----------------------------------------------

-   Addition and subtraction: Commensurability
-   Multiplication and division
-   Units in exponents, sines, logs, etc?

### Units for proportionality constants in models

-   Spring constant k in Hooke's law \[ F = k x \]
-   Newton's law of cooling \[ dT/ dt = k (T-T_ambient) \]

### Calculus and units

Arithmetic of units suggests the appropriate units for derivatives and integrals.

Attention to: relative rates and elasticity.

Examples with derivatives: Temperature gradients (Temp per length), concentration gradients, pressure gradients, velocity (Length per time), flow rates (Volume per time).

Examples with integrals: man-hours, kilowatt-hours, growing degree days, work.

Unit conversion
------------------------------

What are the units of body-mass index, e.g., the one where 25 is the upper limit of normal weight, and 30 and above is obese? (Note: mass per height squared -- kg / m^2 ) That's why there is a conversion factor when height is measured in inches and weight in pounds. Why is it based on square length rather than length cubed (which would be proportional to mass in a simple model)? Because the output should be like a length -- how fat you are, e.g., waist circumference. (Maybe put this in dimension conversion, since a density would let you convert volume to kg.

Conversion between commensurates, e.g., dollars to yen, meters to miles, cups to liters.

Example problems. Air pressure at sea level is roughly 10.14 Newtons per square centimeter. Convert this to pounds per square inch.

**Ans.** 1 Newton is 0.2248 pounds, and 1 sq. cm is 0.155 square inches. Plugging these values into (10.14 Newtons)/(1 sq cm) gives (10.14 \* 0.2248 pounds)/(0.155 sq inches) = 14.706 pounds per square inch.

Dimension Conversion
---------------------------------------------

Unlike unit conversions, which are automatic and unambiguous, a dimension conversion involves some auxiliary information. A simple, everyday dimension conversion occurs with travel, e.g., you're walking at 3 miles per hour and you have 2 miles to go; how long will this take.

Commonly encountered situations (particularly, I think, those that are analogous to velocity and distance)

-   Velocity to distance
-   acceleration to velocity
-   Wages to pay: e.g., dollars per hour into a weekly paycheck
-   Power to energy (e.g., kwh to watts)

Example: Air pressure should be given as a pressure, that is, force per unit area. However, there are various traditional units, and metric units sometimes follow the traditional units.

Air pressure at sea level is about 10.14 Newtons per sq. cm. That's force per unit area. Those used to the traditional units can convert the metric units to find the air pressure in pounds per square inch, which is also a force per unit area.

However, air pressure is sometimes given in these units: inches of mercury (written "inches Hg"), mm Hg, or kg/cm^2. None of these are a force per unit area, but there is a dimension conversion.

In these various units, air pressure at sea level is given as 760 mm Hg or 29.92 inches Hg or 1.033 kg/sq cm.

1.  A kilogram (that is, kg) is a unit of mass, not force. But given that we live at the surface of the Earth, it's natural to think of the force exerted by 1 kg due to gravity. The force is mass times acceleration. Look up a value for acceleration due to gravity at the Earth's surface, and use this to do a dimension conversion of 1.033 kg/sq cm to Newtons/sq. cm.
2.  The measurement of pressure in terms of a length of mercury comes from the idea that a column of mercury of a given height will push down with a given pressure. Find the mass density of mercury given that 760 mm Hg will produce a pressure of 10.14 Newtons per sq cm. (Remember that force is mass times acceleration, and the relevant acceleration is that due to gravity at the Earth's surface.)


### The Units of Information

Explain the idea of information and the common unit, bits. It's related to the use of log-base-2 in the definition of information.

### For fun

-   <http://www.unc.edu/~rowlett/units/>
-   A recording of a units-based billing dispute: Is a penny different from a dollar? [See the recording on this page.](http://verizonmath.blogspot.com/2007/08/original-recording-of-verizon-customer.html) or [for more of the story](http://verizonmath.blogspot.com/2006/12/verizon-doesnt-know-dollars-from-cents.html)
    .
    
### and more
    
-   Start units and dimensions: Given a list of words students put them into natural groups. The words were units (foot, liter, acre, teaspoon, meter, hour, hectare, day, etc). The groups were dimensions. Sole goal: to distinguish between units and dimensions.
-   Dimensions are what you measure, units (of measurement) are the scale by which you measure them.
-   Dimensions: Class made a list of dimensions: length, time, mass, velocity, force, angle, etc. Systematic discussion of dimensions for next class period
-   Units: class proposed 2 units for each dimension.
-   Different units for same dimension are related multiplicatively: can convert. The relation can be expressed three ways: (1) 1 foot = 12 inch; (2) foot/inch = 12; (3) (1 foot)/(12 inch) = 1. Note that (2) illustrates the fact that a dimensionless unit (such as foot/inch) is actually a number. Note that (3) (which gives a "flavor of 1") is the useful form for conversion from one unit to another.
-   Example: How many seconds in a year: done by multiplying 1 year by "flavors of 1" such as (365 day)/(1 year) so that units cancel: 1 year = (1 year)x(365 day)/(1 year)x (24 hour)/(1 day) x (60 minute)/(1 hour) x (60 sec)/(1 minute) = 365 x 24 x 60 x 60 second.

### and more still

#### Starting Units and Dimensions

Dimension is the what, unit is the size, value is how many of that size.

-   Physical dimensions: length, area, volume, velocity, acceleration, force , pressure, density, energy
    -   Group the names of units by dimension: mile, meter, mph, acre,acre-foot, cup, bushel, newton, pascal, ...
    -   Work through the dimensions of physical quantities: force, pressure, density
    -   Include money as a dimension: M, L, T, \$
    -   Walk through class: Give me a unit for "mass *(kg, slug), "force* (newton, pound), "energy *(kwh, BTU, therm),*
-   Arithmetic on units. Addition/Subtraction, Multiplication/Division, Exponentiation, Logs/Sines/etc.
-   Table of physical quantities and their dimension [Table 7.1 Giordano and
    Weir](http://dl.dropbox.com/u/5098197/Math135/Figures/Dimension-Table-7.1-Giordano-Weir.pdf)
    -   Transcendental functions MUST take a dimensionless and unitless argument: exp, sin, log, ...
    -   Special nature of log. $\ln(x/x_0)$ has the right form. But what happens if you leave off $x_0$? It's a shifting factor. If you are looking at differences of logs, the shifts cancel out.
-   Conversion between units of the same dimensions. The idea of "flavors of 1"
-   Elasticities and the desire to compare things in different units but in a unitless way.

Example:

-   Convert meters per second to miles per hour. What's the equivalent of 3 m/s. (This is a speed the instructor can run across the room.)
-   Convert pounds per square inch (psi) to newtons per square meter (pascal). Context, bicycle tire pressure. What's the equivalent of 80 psi
-   Convert mmHg to pounds per square inch. (This requires knowing the density of mercury.)

Emphasize the importance of checking the reasonability of answers and the ease with which one can mistakenly invert a flavor of 1. For instance, if you invert meters per mile, you change the calculated "speed" from about 7 mph to more than a million!

#### Quick Units Questions

-   How many days in a month? In some calendars, it's fixed: 28 days.
    -   In the Jewish calendar, the number of days in a month is fixed, but the number of months in a year varies.
    -   In the Arabic calendar, both the number of days in a month and the number of months in the year are fixed. But the calendar moves around the astronomical year.
    -   The Gregorian calendar (promulgated by Pope Gregory, so we can call it a Catholic calendar, and indeed it was adopted in Catholic countries like Spain and Portugal about 200 years before it was adopted by the protestant countries (1754 in England and the colonies), and was not adopted in Eastern Orthodox countries until about 100 years ago (which is why the October revolution happened in November)) has a fractional number of days on average and requires switching between years of length 365 and 366.
-   The "zero" year is arbitrary, like zero temperature in celsius and fahrenheit (and even Kelvin and rankine, although absolute zero has a clear meaning in terms of the natural gas law $PV = nRT$ -- Remember, temperature is really the partial derivative of entropy with respect to internal energy.) In the Jewish year, tonight is the start of 5772, in the Chinese calendar it's 4707, in the Arabic calendar it's 1432. In the Mayan calendar, it's hard to figure out ... "Misinterpretation of the Mesoamerican Long Count calendar is the basis for a New Age belief that a cataclysm will take place on December 21, 2012. December 21, 2012 is simply the day that the calendar will go to the next b'ak'tun."
    -   Question: When was New Year's Day when and where George Washington was born? March 16. The calendar went from March 15, 1732 to the next day, March 16, 1733.
    -   Question: When does a day start? In the Gregorian system, at midnight. In the Jewish system, at local sunset. In the British Royal Navy, at local noon.


```{r pine-lose-clock-1, echo = FALSE}

```

```{r pine-lose-clock-2, echo = FALSE}

```
