---
chapter: "none yet"
author: "Daniel Kaplan"
date: 2019-09-18 
version: 0.1
tags: [notes, dimension, dimensional analysis]
id: spruce-bet-dress 
---

(ref:spruce-bet-dress) Exercise in file spruce-bet-dress

```{r spruce-bet-dress-default, include = FALSE}
SDSdata::sds_setup()
```

TITLE GOES HERE: Dimensional analysis

Example: Period of pendulum.

-   Quantitles of importancePeriod $T$, Pendulum length
    $L$, Mass $M$, Gravity
    $L/T^2$, Angle (dimensionless).
    Combine these together
    $T^a \ L^b \ M^c \ (L/T^2)^d $
    Find $a,b,c,d$ to give a dimensionless quantity. This
    amounts to
    $\mbox{Period}^2 \frac{\mbox{gravity}}{\mbox{length}} =
    f(\mbox{others})$
    Hold the angle constant and solve, giving $\mbox{Period} =

\sqrt{\frac{\mbox{Length}}{\mbox{gravity}}}$

DO THE ESTIMATION OF TIME BY SWINGING THE PENDULUM. At a fixed angle of release, find the period. Change the mass and the length. Make a table of the period versus length. See if changing the mass changes anything. See if things change from very small angles to very large angles. 
Example: Distance travelled by a projectile.

\#\# Friday Sept 23

Dimensions

-   Dimensions: Basic dimensions in science and geometry are length L,
    time T, mass M. Class expressed preceding dimensions (after deleting
    money etc) in terms of L, T, M. Notice that angle is L^0 T^0 M^0
    = 1, referred to (perhaps misleadingly) as dimensionless.
-   Explicit discussion of Work, Energy, Power.
-   All terms on both sides of an equation must have the same dimension.
    Used to advantage in two examples:
-   Example 1: period P of a pendulum. Modeled based on l = pendulum
    length, g = acceleration of gravity, m = mass of pendulum bob.
    Result: P = k \sqrt{l/g}. Note we did not need mass. Seems like
    magic, but the real depth, intuition came in in the choice of
    modeling parameters l, g, m. What if you needed more - for example a
    viscosity parameter or mass density parameter for swinging in a
    liquid, etc.
-   Example 2: range R of a projectile: modeling parameters muzzle
    velocity v, acceleration of gravity g, mass m, initial angle theta.
    R = f(theta) v^2/g. The unknown function of theta comes in because
    theta is dimensionless, we can not fix a power of theta. f(theta)
    determined by more theory or by experiment.
-   QUIZ 2
    -   Read contour diagram
    -   Linear function of 2 variables, from data
    -   Unit conversion

#### Kaplan

-   News from the world of physics: [faster than light?](http://www.nytimes.com/aponline/2011/09/22/world/europe/AP-EU-Breaking-Light-Speed.html?pagewanted=1&hp)
    
-   Seque into the dimensional analysis of $E = m c^2$. What else could it be?

##### Dimensions

Does $x_0 + v_0 t + 1/2 a t^2$ make sense dimensionally?

Calculate my body mass index. Convert mass from pounds to kg (on earth), height from inches to meters.

What are the dimensions of the Body Mass Index? Must there be units to it. Show my BMI in kg/m2 must be equal to my BMI in lbs/in2, and what the conversion factor must be.

Units of coefficients. Work through coefficients on $a x^2 + b x + c$ when $x$ has units of meters. Does the quadratic formula make sense dimensionally?

##### Dimensional Analysis

Pendulum example. What might the period depend on: mass, length, gravity, angle (also, air density, ...) 

Find two dimensionless groups (angle and gravity * time^2^ / length). These must be functions of one another. Work through the implications to get $\sqrt{L}$ dependence of period. Test this.


Construct quantities that are dimensionless. The variables involved are related to one another accordingly. Two examples:

1.  Distance travelled by a canonball. Quantities of interest: velocity v, angle alpha, gravity g, mass m, distance d. How to get a quantity of dimension L by combining quantities of dimension $v \rightarrow L/T$, $g \rightarrow L/T^2$, and $m \rightarrow M$? It's $v^2 / g$. We've just discovered about energy going as $v^2$.
2.  Swing a rock on a string in a circle. What's the force on the string? (The so-called centripetal force.) Force has dimension \$f \rightarrow M L/T^2\$. The rock has mass $m\rightarrow M$ and the string has radius $r \rightarrow L$. Two ways to think about it:
    -   What's the linear velocity of the rock? $v \rightarrow L/T$
    -   What's the angular velocity (revolutions per second)? $ \omega \rightarrow 1/T$
 
How to combine linear velocity, mass, and radius to get $M L/T^2$? It's $m v^2 / r$. Or, for angular velocity, it's $m r / \omega^2$. Compare those to the formulas in a physics book.

### ... and more



Worked on the "corn energy efficiency" problem.

Introduce the Atomic Bomb group project.

```{r spruce-bet-dress-1, echo = FALSE}

```

```{r spruce-bet-dress-2, echo = FALSE}

```
