---
chapter: "MMAC 1.1.81"
title: "Title for doe-stand-magnet"
author: "Danny Kaplan"
difficulty: "XX"
output: learnr::tutorial
tutorial:
  id: "doe-stand-magnet"
  version: 0.1
runtime: shiny_prerendered
date: 2020-07-06 
tags: [MMAC, 1.1.81, third]
id: doe-stand-magnet
---

```{r doe-stand-magnet-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(mosaic)
library(mosaicCalc)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` Using R, define the function

$$f(x) \equiv \frac{\pi(9.852 x^3 - 10.375 x)}{(9 \pi)^2 + 27 x}$$

To help you get started, we've put some text in the code box. But the text is not correct R syntax. You'll have to fix it so that it is a legitimate R expression. Then you'll have to make sure that your expression says the same thing as the mathematical definition.
(Hint: The spacing of the tilde-expression across lines helps to make it easier to read.)

```{r doe-stand-magnet-E1-setup}
# Setup code goes here
```

```{r doe-stand-magnet-E1, exercise=TRUE, exercise.cap="Define f()", exercise.setup="oe-stand-magnet-E1-setup", exercise.eval=FALSE, exercise.lines=4, exercise.completion=FALSE, exercise.startover=TRUE, exercise.diagnostics=TRUE}
f <- makeFun(
  pi(9.852x3 - 10.375x) / 9 pi 2 + 27 x ~ x
)
```


```{r doe-stand-magnet-E1-check}
f <- makeFun(
  pi*(9.852*x^3 - 10.375*x) / ((9 * pi)^2 + 27 * x) ~ x
)
grade_result(
  fail_if( ~ !inherits(.result, "function"), "You should be creating a function"),
  fail_if( ~ !names(formals(.result)) == "x", "The argument to the function should be named `x`."),
  fail_if( ~ .result(10) != f(10),
           "Your tilde expression doesn't exactly correspond to the mathematical definition given in the problem."),
  pass_if( ~ TRUE)
)
```



