---
chapter: "none yet"
author: "Daniel Kaplan"
difficulty: "XX"
date: 2020-09-21 
version: 0.1
tags: [first, second, third]
id: dolphin-toss-car
---

```{r dolphin-toss-car-default, include = FALSE}
library(etude)
library(mosaic)
library(mosaicCalc)

```

`r etude::exercise_title()` Here's a contour diagram showing the depth of a lake and the landscape around it.  **Negative values** are underwater. 

```{r one, echo = FALSE, warning=FALSE, message=FALSE}
f = rfun( seed=78344 )
ff = function(x,y){-5*(f(x/100, y/100)-3)}
contour_plot(ff(x, y) ~ x + y, domain(x=c(-700,700), y=c(-700,700)), filled=FALSE,
             contours_at=c(-20,-15,-10,-5,0, 5,10,15), skip = 0) %>%
  contour_plot(ff(x, y) ~ x + y, contours_at=0, size=4, filled=FALSE)
```


The axis scales and contour units are in feet.

For each of the following, mark on the graph a location where   the conditions apply.  Make sure to use the letters, A -- D to   distinguish the various points from one another.

1. The lake bottom is **steep** and fairly **deep**.
#. The lake bottom is **steep** and **shallow**.
#. The lake bottom is **relatively flat** and **deep**.
#. The lake bottom is **relatively flat** and **shallow**.

