---
chapter: "none yet"
title: "Title for fly-fall-table"
author: "Daniel Kaplan"
difficulty: "XX"
output: learnr::tutorial
tutorial:
  id: "fly-fall-table"
  version: 0.1
runtime: shiny_prerendered
date: 2020-09-22 
tags: [first, second, third]
id: fly-fall-table
---

```{r fly-fall-table-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(ggformula)
library(mosaic)
library(mosaicCalc)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` 

```{r echo=FALSE}
set.seed(133)
pts <- tibble(x = -5 +  cumsum(0.5 + rexp(5) ), y = runif(5))
refx <- sort(c(-3.7, -2.4, -0.95, 3.62,  2.47 ))
f <- spliner(y ~ x, data =  pts)
special_pts <- tibble(x = refx, y = 8 * f(refx) - 3)
label <- character(8)
special_pts$label = 
  sapply(1:length(refx),
        FUN = function(x) paste0("x", x))
gf_fun(8*f(x)-3 ~ x, xlim = c(-4, 4)) %>%
  gf_labs(y = "f(x)") %>%
  
  gf_point(y ~ x, data = special_pts) %>%
  gf_segment(15 + y ~ x + x, data = special_pts, linetype = 2)  %>%
  gf_label(15 ~ x, label = ~ label, data = special_pts )
```

1. At which point is the first derivative positive and the second derivative positive as well? {$x_1$, $x_2$, $x_3$, *$x_4$*, none}
#. At which point is the first derivative *negative* and the second derivative positive? {$x_1$, $x_2$, $x_3$, $x_4$, *none*}
#. At which point is the first derivative positive and the second derivative negative? {*$x_1$*, $x_2$, $x_3$, $x_4$, none}
#. At which point is the first derivative negative and the second derivative negative as well? {$x_1$, *$x_2$*, $x_3$, $x_4$, none}
#. At which point is the second derivative closest to zero?  {$x_1$, $x_2$, *$x_3$*, $x_4$}



