---
chapter: "Derivatives"
author: "Daniel Kaplan"
difficulty: "XX"
date: 2020-09-22 
version: 0.1
tags: [derivatives, average, population]
id: frog-lose-car
---

```{r frog-lose-car-default, include = FALSE}
library(etude)
library(tibble)
library(ggformula)
```

`r etude::exercise_title()` 

```{r echo=FALSE}
Table <- tribble(
  ~ year, ~ population,
  1975, 4.08,
  1980, 4.45,
  1985, 4.84,
  1990, 5.27,
  1995, 5.68,
  2000, 6.07,
  2005, 6.45
)
knitr::kable(Table, caption = " World population, in billions")
gf_point(population ~ year, data = Table) %>%
  gf_line() %>%
  gf_text(population ~ year, label = ~ paste(population, "B", sep=""))
```

#. What are the units of average rate of change for the data below?
    a. Years
    #. Billions of people
    #. *Billions of people per year*
    #. Years per billions of people
    #. Years $\times$ billions of people
    
#. Using the data, find the average rate of change in the world's population between 1975 and 2005.
    a. $6.45/2005$
    #. $4.08/1975$
    #. $1975/4.08$
    #. $2005/6.45$
    #. $\frac{4.08 - 6.45}{2005-1975}$
    #. *$\frac{6.45 - 4.08}{2005-1975}$*
    #. $\frac{2005-1975}{4.08 - 6.45}$
    #. $\frac{2005-1975}{6.45 - 4.08}$

#. Which is the best estimate of the derivative of the population curve in $2005$?
     a. $2005/6.45$
     #. $6.45/2005$
     #. *$0.38/5$*
     #. $2.37/30$


