---
chapter: "Derivatives"
title: "Title for goldfish-speak-blanket"
author: "Daniel Kaplan"
difficulty: "XX"
output: learnr::tutorial
tutorial:
  id: "goldfish-speak-blanket"
  version: 0.1
runtime: shiny_prerendered
date: 2020-09-22 
tags: [velocity, acceleration, third]
id: goldfish-speak-blanket
---

```{r goldfish-speak-blanket-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(ggformula)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` Each of the graphs shows the position $x$ of a particle -- I, II,  III, or IV -- as a function of time.

```{r echo=FALSE, out.width = "50%", fig.show="hold"}
gf_fun(t^2 - 3 ~ t, xlim=c(0,3)) %>% 
  gf_labs(title = "Particle I", x="time", y = "position")
gf_fun(sin(2*pi*t/3) - 3 ~ t, xlim=c(0,3)) %>% 
  gf_labs(title = "Particle II", x="time", y = "position")
gf_fun(2 - t/1.5 ~ t, xlim=c(0,3)) %>% 
  gf_labs(title = "Particle III", x="time", y = "position")
gf_fun(sqrt(t) ~ t, xlim=c(0,3)) %>% 
  gf_labs(title = "Particle IV", x="time", y = "position")
```

Note that the vertical scales vary from one graph to another. It may be important to take this into account when answering the following.

1. Which graph shows the particle with constant velocity? {I, II, *III*, IV}
#. Which graph shows the particle with the greatest initial velocity? {I, II, III, *IV*}
#. Which graph shows the particle with positive acceleration throughout?{I, II, III, IV}
#. Which graph shows the particle with zero average velocity?{I, *II*, III, IV}
#. Which graph shows the particle with the greatest average velocity? {*I*, II, III, IV}
#. Which graph shows the particle with zero acceleration? {I, II, *III*, IV}




