---
chapter: "Derivatives"
title: "Title for oak-ring-car"
author: "Daniel Kaplan"
difficulty: "XX"
output: learnr::tutorial
tutorial:
  id: "oak-ring-car"
  version: 0.1
runtime: shiny_prerendered
date: 2020-09-21 
tags: [partial derivs, contour, third]
id: oak-ring-car
---

```{r oak-ring-car-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(mosaic)
library(mosaicCalc)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` Here's a contour plot of a function of two variables:

```{r uvplot3,echo=FALSE, warning=FALSE}
f <- rfun( ~ x | y, seed=672)

Labels <- tibble(x=c(-1.2,0,1.2,-1.2,0,1.2,-1.2,0,1.2),
                 y=c(1,1,1,0,0,0,-1,-1,-1),
                 label = toupper(c("a","b","c","d","e","f","g","h","i")))
contour_plot(f(u,v)~  u +  v, 
             domain(u=range(-2,2), v=range(-2,2) ),
             skip = 0) %>%
  gf_label(y ~ x, label = ~ label, data = Labels)
```
 

At each of the marked points, say whether the partial derivative is positive, negative, or zero.  (NOTE: It's hard to say from a contour graph like this when something is exactly zero.  And if it's not exactly zero, then it must be either positive or negative.  Here, use "zero" to mean "much smaller in magnitude than at most of the other points.)

```{r oak-ring-car-QA1, echo=FALSE}
etude::choose_one(
  prompt = "Point A. $\\partial f/\\partial u$ and $\\partial f/\\partial u$?",
  choices = c(
    "pos & pos", "pos & zero", "pos & neg",
    "zero & pos", "zero & zero", "zero & neg",
    "neg & pos", "+neg & zero+", "neg & neg"
  ), random_answer_order = FALSE, inline = FALSE
)
```



```{r oak-ring-car-QA2, echo=FALSE}
etude::choose_one(
  prompt = "Point B. $\\partial f/\\partial u$ and $\\partial f/\\partial u$?",
  choices = c(
    "pos & pos", "pos & zero", "pos & neg",
    "zero & pos", "zero & zero", "zero & neg",
    "neg & pos", "+neg & zero+", "neg & neg"
  ), random_answer_order = FALSE, inline = FALSE
)
```

```{r oak-ring-car-QA3, echo=FALSE}
etude::choose_one(
  prompt = "Point C. $\\partial f/\\partial u$ and $\\partial f/\\partial u$?",
  choices = c(
    "pos & pos", "pos & zero", "+pos & neg+",
    "zero & pos", "zero & zero", "zero & neg",
    "neg & pos", "neg & zero", "neg & neg"
  ), random_answer_order = FALSE, inline = FALSE
)
```

```{r oak-ring-car-QA4, echo=FALSE}
etude::choose_one(
  prompt = "Point D. $\\partial f/\\partial u$ and $\\partial f/\\partial u$?",
  choices = c(
    "pos & pos", "pos & zero", "pos & neg",
    "zero & pos", "zero & zero", "zero & neg",
    "neg & pos", "neg & zero", "+neg & neg+"
  ), random_answer_order = FALSE, inline = FALSE
)
```

```{r oak-ring-car-QA5, echo=FALSE}
etude::choose_one(
  prompt = "Point E. $\\partial f/\\partial u$ and $\\partial f/\\partial u$?",
  choices = c(
    "pos & pos", "pos & zero", "pos & neg",
    "zero & pos", "zero & zero", "zero & neg",
    "neg & pos", "neg & zero", "+neg & neg+"
  ), random_answer_order = FALSE, inline = FALSE
)
```


```{r oak-ring-car-QA6, echo=FALSE}
etude::choose_one(
  prompt = "Point F. $\\partial f/\\partial u$ and $\\partial f/\\partial u$?",
  choices = c(
    "pos & pos", "pos & zero", "+pos & neg+",
    "zero & pos", "zero & zero", "zero & neg",
    "neg & pos", "neg & zero", "neg & neg"
  ), random_answer_order = FALSE, inline = FALSE
)
```

```{r oak-ring-car-QA7, echo=FALSE}
etude::choose_one(
  prompt = "Point G. $\\partial f/\\partial u$ and $\\partial f/\\partial u$?",
  choices = c(
    "pos & pos", "pos & zero", "pos & neg",
    "zero & pos", "zero & zero", "+zero & neg+",
    "neg & pos", "neg & zero", "neg & neg"
  ), random_answer_order = FALSE, inline = FALSE
)
```


```{r oak-ring-car-QA8, echo=FALSE}
etude::choose_one(
  prompt = "Point H. $\\partial f/\\partial u$ and $\\partial f/\\partial u$?",
  choices = c(
    "pos & pos", "pos & zero", "pos & neg",
    "zero & pos", "zero & zero", "zero & neg",
    "neg & pos", "neg & zero", "+neg & neg+"
  ), random_answer_order = FALSE, inline = FALSE
)
```

