---
chapter: "derivatives"
title: "Title for pine-lead-car"
author: "Daniel Kaplan"
difficulty: "XX"
output: learnr::tutorial
tutorial:
  id: "pine-lead-car"
  version: 0.1
runtime: shiny_prerendered
date: 2020-10-12 
tags: [derivatives, monotonicity]
id: pine-lead-car
---

```{r pine-lead-car-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```

```{r echo=FALSE}
source(system.file("Exercises/Scripts/function_set_A.R", package = "math141Zexercises"))
```


`r etude::exercise_title()` Each of the following graphs shows the derivative of some function $f()$. (Note: the graph doesn't show $f()$ but rather the function $f()$ which is the derivative of $f()$.) For each graph, say whether the function $f()$ is monotonically increasing, monotonically decreasing, or neither. (Note that the horizontal scale is the same in every graph, but the vertical scale can be different from one scale to another.)

```{r echo=FALSE, out.width = "50%", fig.show = "hold"}
slice_plot(df1(x) ~ x, domain(x = c(-5,5))) %>% gf_labs(title="Derivative of function A", y="df(x)")
```

```{r pine-lead-car-QA1, echo=FALSE}
etude::choose_one(
  prompt = "Function A is ...",
  choices = list(
    "+monotonically increasing+" = "A monotonically increasing function has a derivative function that is everywhere $> 0$",
    "monotonically decreasing" = "A monotonically decreasing function has a derivative function that is everywhere $< 0$",
    "constant" = "A constant function has a derivative that is everywhere 0.",
    "non-monotonic" = "A non-monotonic function goes up and down, hence the derivative is positive in some places and negative in others.",
    "Can't tell from the info provided" = "This is the case if you cannot tell if the derivative is positive or negative."
  ), 
  random_answer_order = FALSE,
  allow_retry = FALSE
  
)
```

```{r echo=FALSE, out.width = "50%", fig.show = "hold"}
slice_plot(df2(x) ~ x, domain(x = c(-5,5))) %>% gf_labs(title="Derivative of function B", y="df(x)")
```

```{r pine-lead-car-QA2, echo=FALSE}
etude::choose_one(
  prompt = "Function B is ...",
  choices = list(
    "monotonically increasing" = "A monotonically increasing function has a derivative function that is everywhere $> 0$",
    "+monotonically decreasing+" = "A monotonically decreasing function has a derivative function that is everywhere $< 0$",
    "constant" = "A constant function has a derivative that is everywhere 0.",
    "non-monotonic" = "A non-monotonic function goes up and down, hence the derivative is positive in some places and negative in others.",
    "Can't tell from the info provided" = "This is the case if you cannot tell if the derivative is positive or negative."
  ), 
  random_answer_order = FALSE,
  allow_retry = FALSE
  
)
```

```{r echo=FALSE, out.width = "50%", fig.show = "hold"}
slice_plot(df3(x) ~ x, domain(x = c(-5,5))) %>% gf_labs(title="Derivative of function C", y="df(x)")
```


```{r pine-lead-car-QA3, echo=FALSE}
etude::choose_one(
  prompt = "Function C is ...",
  choices = list(
    "monotonically increasing" = "A monotonically increasing function has a derivative function that is everywhere $> 0$",
    "monotonically decreasing" = "A monotonically decreasing function has a derivative function that is everywhere $< 0$",
    "constant" = "A constant function has a derivative that is everywhere 0.",
    "+non-monotonic+" = "A non-monotonic function goes up and down, hence the derivative is positive in some places and negative in others.",
    "Can't tell from the info provided" = "This is the case if you cannot tell if the derivative is positive or negative."
  ), 
  random_answer_order = FALSE,
  allow_retry = FALSE
  
)
```

```{r echo=FALSE, out.width = "50%", fig.show = "hold"}
slice_plot(df4(x) ~ x, domain(x = c(-5,5))) %>% gf_labs(title="Derivative of function D", y="df(x)")
```

```{r pine-lead-car-QA4, echo=FALSE}
etude::choose_one(
  prompt = "Function D is ...",
  choices = list(
    "+monotonically increasing+" = "A monotonically increasing function has a derivative function that is everywhere $> 0$",
    "monotonically decreasing" = "A monotonically decreasing function has a derivative function that is everywhere $< 0$",
    "constant" = "A constant function has a derivative that is everywhere 0.",
    "non-monotonic" = "A non-monotonic function goes up and down, hence the derivative is positive in some places and negative in others.",
    "Can't tell from the info provided" = "This is the case if you cannot tell if the derivative is positive or negative."
  ), 
  random_answer_order = FALSE,
  allow_retry = FALSE
  
)
```

```{r echo=FALSE, out.width = "50%", fig.show = "hold"}
slice_plot(df5(x) ~ x, domain(x = c(-5,5))) %>% gf_labs(title="Derivative of function E", y="df(x)") %>%
  gf_theme(axis.text.y=element_blank())
```


```{r pine-lead-car-QA5, echo=FALSE}
etude::choose_one(
  prompt = "Function E is ...",
  choices = list(
    "monotonically increasing" = "A monotonically increasing function has a derivative function that is everywhere $> 0$",
    "monotonically decreasing" = "A monotonically decreasing function has a derivative function that is everywhere $< 0$",
    "constant" = "A constant function has a derivative that is everywhere 0.",
    "non-monotonic" = "A non-monotonic function goes up and down, hence the derivative is positive in some places and negative in others.",
    "+Can't tell from the info provided+" = "This is the case if you cannot tell if the derivative is positive or negative."
  ), 
  random_answer_order = FALSE,
  allow_retry = FALSE
  
)
```

```{r echo=FALSE, out.width = "50%", fig.show = "hold"}
slice_plot(df6(x) ~ x, domain(x = c(-5,5))) %>% gf_labs(title="Derivative of function F", y="df(x)")
```

```{r pine-lead-car-QA6, echo=FALSE}
etude::choose_one(
  prompt = "Function F is ...",
  choices = list(
    "+monotonically increasing+" = "A monotonically increasing function has a derivative function that is everywhere $> 0$",
    "monotonically decreasing" = "A monotonically decreasing function has a derivative function that is everywhere $< 0$",
    "constant" = "A constant function has a derivative that is everywhere 0.",
    "non-monotonic" = "A non-monotonic function goes up and down, hence the derivative is positive in some places and negative in others.",
    "Can't tell from the info provided" = "This is the case if you cannot tell if the derivative is positive or negative."
  ), 
  random_answer_order = FALSE,
  allow_retry = FALSE
  
)
```
