---
chapter: "Marginal vs average"
title: "Title for rabbit-hear-car"
author: "Daniel Kaplan"
difficulty: "XX"
output: learnr::tutorial
tutorial:
  id: "rabbit-hear-car"
  version: 0.1
runtime: shiny_prerendered
date: 2020-09-22 
tags: [marginal vs average, second, third]
id: rabbit-hear-car
---

```{r rabbit-hear-car-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(tibble)
library(ggformula)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` The graph comes from the classical economics of the theory of the firm. It compares the total cost of producing a quantity $q$ of a product to the total revenue generated by selling that quantity on the market.

```{r echo = FALSE, warning=FALSE}
data <- tribble(
  ~ quantity, ~ money, ~ which, ~label,
  5,  5*130/25, "revenue", NA,
  25, 130, "revenue", "R",
  5, 60, "cost", NA,
  25, 100, "cost", "C"
)
gf_line(money ~ quantity , color = ~ which, data = data) %>%
  gf_lims(x = c(0,30)) %>%
  gf_text(money ~ quantity, label = ~label, color = "black", xnudge=3)
```

a. At which quantity is the company losing money? {*10*, 16, 20}
#. At quantity = 10, which is greater: the marginal cost of production or the marginal revenue from sales? {marginal cost, *marginal revenue*, they are the same}
#. At quantity = 16, which is greater: the marginal cost of production or the marginal revenue from sales? {marginal cost, *marginal revenue*, they are the same}
#. Assuming that all the sales are made at the same price (or, more precisely, net revenue), what is the price of one unit? Choose the closest. {$2 per unit, *$4 per unit*, $6 per unit }



