---
chapter: "none yet"
title: "Title for birch-sell-saucer"
author: "Danny Kaplan"
output: learnr::tutorial
tutorial:
  id: "birch-sell-saucer"
  version: 0.1
runtime: shiny_prerendered
date: 2020-06-05 
tags: [fermi problem, pendulum, rock climbing]
id: birch-sell-saucer
---

```{r birch-sell-saucer-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(ggformula)
library(mosaic)
library(mosaicCalc)
library(math141Z)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` At the very beginning of [this 1987 video](http://www.ukclimbing.com/news/item.php?id=49981), the climber, [Catherine Destivelle](https://en.wikipedia.org/wiki/Catherine_Destivelle), is dangling from a swinging rope. Make an estimate of how long the rope is, based on principles of dimensional analysis. (Hint: You might have to do an measurement with a short string and a small mass in your room. No rock climbing required or recommended!)

```{r birch-sell-saucer-Essay1}
etude::essay_response(
  "Give your estimate of the length of the rope, explaining the reasoning that went into it."
)
```

PS. Glad to report that as of July 2020, Catherine is 60 years old.

