---
chapter: "Benjamin Franklin and oil"
title: "Title for doe-fly-fridge"
author: "Danny Kaplan"
difficulty: "XX"
output: learnr::tutorial
tutorial:
  id: "doe-fly-fridge"
  version: 0.1
runtime: shiny_prerendered
date: 2020-07-11 
tags: [first, second, third]
id: doe-fly-fridge
---

```{r doe-fly-fridge-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
show_answers(TRUE)
knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` Benjamin Franklin (Publisher, scientist, and American statesman, 1706-1790) put one teaspoon of oil on Clapham Common Pond on a windy day to see how it reduced the size of waves.  He observed that the oil spread out to cover about one-half acre.  

In Franklin's time, no one knew about molecules.  We now know that such oil spreads out in a layer that is one molecule thick.  


```{r doe-fly-fridge-sandbox1, exercise=TRUE, exercise.cap="Sandbox",  exercise.eval=FALSE, exercise.lines=2, exercise.completion=FALSE, exercise.startover=TRUE, exercise.diagnostics=TRUE}

```
```{r doe-fly-fridge-QA1, echo=FALSE}
etude::choose_one(
  prompt = "1) Based on Franklin's measurements, how large is an oil molecule in nanometers. (A nanometer is $10^{-9}$ meters.)",
  choices = c("0.25 nm", "+2.5 nm+", "25 nm", "250 nm", "2500 nm"),
  random_answer_order = FALSE
)
```



```{r doe-fly-fridge-QA2, echo=FALSE}
etude::choose_one(
  prompt = "2) How does this compare to modern estimates of the size of an oil molecule?",
  choices = c("Franklin's estimate was much too big.",
              "+Franklin's estimate was just about right.+",
              "Franklin's estimate was much too small."),
  random_answer_order = FALSE,
  inline = FALSE
)
```


````{etude}
Half an acre is 2023 square meters or 20230000 square cm.

One teaspoon is 4.93 ml, that is, 4.93 cc.  

The area times depth will give the volume, so 20230000*d = 4.93, so the depth in cm is `4.93 / 20230000`.


One nm is $10^7$ nm, so the depth of the oil layer is `1e7*4.93 / 20230000`.

````

[Motivated by ``Measuring Molecules'' in Harte (1985) \cite{harte-1985}.]


