---
title: "Differentiation exercises"
author: "Daniel Kaplan"
subject: "derivatives"
---

```{r include=FALSE}
library(mosaic)
library(mosaicCalc)
```

# Exercise 6: Cobb-Douglas

The "Cobb-Douglas production function" is a simple mathematical model how labor $L$  and capital $K$ combine to produce output $P$. It is
$$ P(L,K) = A K^\alpha L^{1-\alpha} .$$

For simplicity, imagine that capital and labor are both measured in dollars per year --- the amount that the labor force is paid in a year and the amount that one could rent a factory for a year. 

a. If production $P(L,K)$ is also measured in dollars per year   (say, the value of the factory output), what is the dimension of the   constant $A$?
  
#.  According to the model, what happens to production if both $K$ and $L$ are increased by   a factor constant factor $\beta$?  (Hint: Substitute in $K \rightarrow \beta K$ and $L \rightarrow \beta L$ and simplify.)

#.   Consider a particular factory with $A=2.5$ and $\alpha=0.33$. In R, implement the function $P(K,L)$.  Use your function to compute   the production of the factory for $K=10$ and $L=20$.  Confirm that   you get $P(K=10,L=20) = 39.78$.
  
<!-- begin ANSWER -->
```{r }
P = makeFun( A*(K^a)*(L^(1-a)) ~ K|L, 
                  A=2.5, a=0.33 ) 
P(K=10,L=20)
```
 
<!-- end ANSWER -->

#.  Of course, a factory that rents for $10/yr and where the labor   costs $20/yr is silly.  Calculate the value $P(K,L)$ when $K$ is $10 million/yr and $L$ is $20 million/yr.


We'll stick with numbers like $K=10$ and $L=20$ to keep things easy to read, but feel free to interpret them as "millions of dollars." 

Congratuations!  Based on your ability to use the Cobb-Douglas model, you've been promoted to manager of the factory.  One of your jobs is to decide how to balance expenditures on capital and labor in order to raise productivity.  

One basic question is what happens when you raise either capital or labor, holding the other one constant.  Using appropriate partial derivatives evaluated at $K=10$, $L=20$, calculate:

a. The rate at which an increase in spending on labor will increase productivity.
  
<!-- begin ANSWER -->
```{r }
dPdL = D( P(K=K,L=L)~L )
dPdL(K=10,L=20)
```
 
<!-- end ANSWER -->
  
#. The rate at which an increase in spending on capital will   increase productivity.
  
<!-- begin ANSWER -->
```{r }
dPdK = D( P(K=K,L=L)~K )
dPdK(K=10,L=20)
```
 
<!-- end ANSWER -->

#. Based on the above, if you had to choose between spending on   capital or labor, and your goal is to increase productivity as much   as possible, which would you spend on, capital or labor?



Your economist friend tells you to watch out for "diminishing marginal returns."  This means that, as you increase spending on either labor or capital, the rate of increase in production tends to diminish.  You'll still get increased production as you increase spending, but it won't increase as fast at high levels of expense as at low levels.


a. Compute the partial derivative of production with respect to   labor at a higher level of labor, say $L=21$, but holding $K=10$.   How does the value of the derivative at $L=21$ compare to that at   $L=20$?  Is this consistent with the idea of "diminishing marginal   returns" for labor?

<!-- begin ANSWER -->
The rate of production increase, when capital spending is increased,
goes down at $L=21$ compared to $L=21$ (holding $K=10$).
```{r }
dPdL(K=10,L=21) - dPdL(K=10,L=20)
```
 
<!-- end ANSWER -->
  
  
#.   Do the same for the partial derivative of production with   respect to capital, evaluated at $L=20$ and $K=11$.  How does the   value of the derivative at $K=11$ compare to that at $K=10$.  Is   this consistent with the idea of "diminishing marginal returns"   for capital?

<!-- begin ANSWER -->
The rate of production increase, when capital spending is increased,
goes down at $K=11$ compared to $K=10$ (holding $L=20$).
```{r }
dPdK(K=11,L=20) - dPdK(K=10,L=20)
```
 
<!-- end ANSWER -->
  
  
#.   Use an appropriate partial second derivative to find the rate of   diminishing partial returns for labor at $L=20$ and $K=10$.  Show   that it's consistent with the difference you got in Part (a).

<!-- begin ANSWER -->
```{r }
dPdL2 = D( P(K=K,L=L)~L|L )
dPdL2(K=10,L=20)
```
 

The second derivative is negative, indicating diminishing marginal returns.  Notice that the magnitude of the second derivative, multiplied by the change by 1 in $L$ from $L=20 \rightarrow 21$, is roughly the same as the answer in (a).

<!-- end ANSWER -->

#.   Use an appropriate partial second derivative to find the rate of   diminishing partial returns for capital at $L=20$ and $K=10$.  Show   that it's consistent with the difference you got in Part (b).

<!-- begin ANSWER -->
```{r }
dPdK2 = D( P(K=K,L=L)~K|K )
dPdK2(K=10,L=20)
```
 
The second derivative is negative, indicating diminishing marginal returns.  Notice that the magnitude of the second derivative, multiplied by the change by 1 in $L$ from $L=20 \rightarrow 21$, is roughly the same as the answer in (b).

<!-- end ANSWER -->

#.  You might think of the rate of increase in production with   respect to labor as the ``value rate" of labor.  Similarly, the rate of   increase in production with respect to capital is the value of   capital.   Due to diminishing marginal returns, an increase in labor   spending, holding capital constant, decreases the value rate of   labor.  Similarly, an increase in capital spending holding labor   spending constant decreases the value rate of capital. 
  
But what happens to the value of labor when capital spending is   increased?  You can answer this by comparing the value rate of labor,   $\frac{\partial  P}{\partial L}$, at two different capital spending   levels, say   $(K=10,L=20)$ and $(K=11,L=20)$.  Notice that even though you're   looking at the rate with respect to labor, you're changing the   expenditure on capital.
  
#. Compare $\frac{\partial P}{\partial L}$ at slightly different     values of $K$, holding $L$ constant at 20.  Does the value rate of     labor increase or decrease with spending on capital?     

<!-- begin ANSWER -->  
```{r }
dPdL(K=11,L=20) - dPdL(K=10,L=20)
```
 
This is a positive change, meaning that increasing spending on capital increases the value of labor.
<!-- end ANSWER -->
    
#. Similarly, compare $\frac{\partial P}{\partial K}$ at slightly different     values of $L$, holding $K$ constant at 20.  Does the value rate of     labor increase or decrease with spending on capital?     

<!-- begin ANSWER -->  
```{r }
dPdK(K=10,L=21) - dPdK(K=10,L=20)
```
 
Similarly, increasing spending on labor, increases the value of capital.  Note that the numerical values are *almost* the same. When we move from a finite difference ($L=20 \rightarrow 21$) to an infinitesimal, the numbers will be exactly the same.

<!-- end ANSWER -->    
    
#.   Finally, construct and evaluate the mixed partial derivative,     $\frac{\partial^2 P}{\partial L \partial K}$ at $K=10, L=20$.     Compare this to the results you got for the  way $\frac{\partial     P}{\partial K}$ changes with increasing $L$ and the way    $\frac{\partial P}{\partial L}$ changes with increasing $K$. 
    
  
<!-- begin ANSWER -->  
```{r }
d2PdLdK = D( P(K=K,L=L)~K|L )
d2PdKdL = D( P(K=K,L=L)~L|K )
d2PdLdK(K=10,L=20) 
d2PdKdL(K=10,L=20) 
```
 
<!-- end ANSWER -->  

The two mixed partials are exactly the same.  As you can see, they are also very close to the rate for the finite difference.    

# Exercise 7


# Exercise 8



     
# Exercise 16

The figure shows a model of farming population as a  function of year:

```{r echo=FALSE}
f <- makeFun(20 + 2000*dnorm(year, mean = 1931, sd =  20) ~ year)
graphFun(f(year) ~  year, yearlim = c(1925, 2015)) %>%
  gf_labs(y = "f(t)")
```
1. If $f^\prime(1950) < 0$, what does this tell you about the farming population in 1950?

    a. The number of farms reached a peak in 1950.
    #. The number of farms is going up in 1950.
    #. *The number of farms is going down 1950.*
    
2. Which of these statements is true?

    a. $f^\prime(1950) > f^\prime(1970)$.
    #. $f^\prime(1950) > f^\prime(1970)$.
    #. *$f^\prime(1950) < f^\prime(1970)$.*
    #. *The number of farms is going down 1950.*
    
# Exercise 12

Consider the following function ...

```{r echo=FALSE}
set.seed(101)
pts <- data_frame(x = -5 +  cumsum(0.5 + rexp(7) ), y = runif(7))
refx <- sort(c(-3.7, -1.66, -0.5, 1.4, 3.62, 1.075, 2.47 ))
f <- spliner(y ~ x, data =  pts)
special_pts <- data_frame(x = refx, y = 8 * f(refx) - 3)
label <- character(8)
special_pts$label = 
  sapply(1:length(refx),
        FUN = function(x) paste0("x", x))
gf_fun(8*f(x)-3 ~ x, xlim = c(-4, 4)) %>%
  gf_labs(y = "f(x)") %>%
  gf_point(y ~ x, data = special_pts) %>%
  gf_segment(3 + y ~ x + x, data = special_pts, linetype = 2)  %>%
  gf_label(3 ~ x, label = ~ label, data = special_pts )
```

1. At which point is $f(x)$ at a global minimum? {$x_3$, *$x_2$*, $x_6$, none of them}
#. At which point is $f(x)$ at a local maximum? {$x_1$, $x_2$, *$x_4$*, $x_7$, none of them}
#. At which point is the derivative (with respect to $x$) the smallest? {*$x_1$*, $x_3$, $x_5$, $x_7$}
#. At which point is the derivative (with respect to $x$) negative? {*$x_1$*, $x_2$, $x_3$, $x_4$, none of them}
#. At which point is the derivative zero? {$x_5$, *$x_6$*, $x_7$, none of them}
#. At which point is the function value zero? {$x_4$, *$x_5$*, $x_6$, none of them}

# Exercise 16

Let $P(t)$ represent the price of a share of stock of a corporation at time $t$.  

1. Consider the statement "the price of the stock today is rising faster and faster." The first and second derivative of $P(t)$ today are, respectively,
    a. $-,\ -$
    #. $-,\ 0$
    #. $-,\ +$
    #. $0,\ -$
    #. $0,\ 0$
    #. $0,\ +$
    #. $+,\ -$
    #. $+,\ 0$
    #. *$+,\ +$*
    
#. Consider the statement "the price of the stock bottomed out last Thursday." The first and second derivative of $P(t)$ last Thursday are, respectively,
    a. $-,\ -$
    #. $-,\ 0$
    #. $-,\ +$
    #. $0,\ -$
    #. $0,\ 0$
    #. *$0,\ +$*
    #. $+,\ -$
    #. $+,\ 0$
    #. $+,\ +$

# Exercise 14




# Exercise 15


# Cost and revenue


# Exercise 18: The gradient tells you about change



## Exercise 19: Mortgage costs


# Exercise 37

Consider the contour diagram shown below of a function $U = f(x, y)$. Say whether each of the given derivatives, at the point $(4, 4)$,  appears to be positive, negative, or zero.

```{r cplot, echo = FALSE}
f = function(x,y){  }
graphFun( 10*(y+1)/(x+1) ~ x + y, 
          xlim=c(0,10), ylim=c(0,10), tile=FALSE) %>% gf_labs(
          levels=c(5,10,15,20,25,30,35,40)) 
```
 
1. $\partial U/\partial x$ 
#. $\partial U/\partial y$
#. $\partial^2 U/\partial x^2$
#. $\partial^2 U/\partial y^2$
#. $\partial^2 U/ (\partial x \partial y)$

# Exercise 38

When constructing a partial derivative, remember this: other than the variable with respect to which differentiation is being taken,  *all other variables* are treated as constants.  This is just a little bit different than ordinary differentiation, where the other variables may well themselves vary with the variable of differentiation.  The fact that a variable is called $y$ doesn't mean that $y$ might not change with $x$.  But when taking a partial derivative $\partial / \partial x$, the variable $y$ will be held constant even as $x$ changes. 

The following examples illustrate this rule.  Make sure you understand them:




Ordinary Derivatives | | Partial Derivatives
----------------|--------|-----------
$\frac{d}{dx} y = \frac{dy}{dx}$  | but |   $\frac{\partial}{\partial x} y = 0$ 
$\frac{d}{dx} x y = y + x \frac{dy}{dx}$  | but |   $\frac{\partial}{\partial x} x y = y$ 
$\frac{d}{dx} \left( x +  y\right) = 1 + \frac{dy}{dx}$  | but |   
$\frac{\partial}{\partial x} \left( x + y \right) = 1$
$\frac{d}{dx} \left( x^2 +  y^2\right) = 2x + 2y\frac{dy}{dx}$  | but |   $\frac{\partial}{\partial x}\left( x^2 + y^2\right) = 2x$. 
\end{tabular}}



Find the following derivatives, being careful to distinguish *partial* derivatives ($\partial$) from ordinary derivatives ($d$).  Following convention, $x$, $y$, $z$, and $t$ represent variables, $a$, $b$, $c$, $a_0$, $a_1$, $b_0$, $b_1$, $\ldots$ are constants, and $f$, $g$ and $h$ are functions.  When the particular form of a function isn't stated, you can write the ordinary derivative (of a function of one variable) as $f'$, $g'$, or $h'$ respectively.  The partial  derivative with respect to $x$ can be written $f_x$, $g_x$, or $h_x$.

Find the symbolic derivatives of these expressions. Pay careful attention to the variable of differentiation; it changes from problem to problem.  Note that some of the problems involve ordinary differentiation, others involve partial differentiation.

1. $\frac{\partial}{\partial x} x$
#. $\frac{\partial}{\partial x} y$
#. $\frac{d}{d x} a x $
#. $\frac{\partial}{\partial x} x y$
#. $\frac{d}{d x} x y$
#. $\frac{\partial}{\partial y} x y$
#. $\frac{d}{d y} x y$
#. $\frac{d}{d y} x y$
#. $\frac{\partial}{\partial x} A e^{kt}$
#. $\frac{\partial}{\partial t} A e^{kt}$
#. $\frac{\partial}{\partial x} A x e^{kt}$
#. $\frac{\partial}{\partial t} A x e^{kt}$
#. $\frac{\partial}{\partial x} \left( a_0 + a_1 x + a_2 x^2\right)$
#. $\frac{\partial}{\partial y} \left( a_0 + a_1 x + a_2 x^2\right)$
#. $\frac{\partial}{\partial x} \left( b_0 + b_1 y + b_2 y^2\right)$
#. $\frac{\partial}{\partial y} \left( a_0 + a_1 x + b_1 y\right)$
#. $\frac{\partial}{\partial x} \left( a_0 + a_1 x + b_1 y\right)$
#. $\frac{\partial}{\partial x} \left( a_0 + a_1 x + b_1 y + c x y\right)$
#. $\frac{\partial}{\partial y} \left( a_0 + a_1 x + b_1 y + c x y\right)$
#. $\frac{\partial}{\partial y} \left( a_0 + a_1 x + b_1 y + c x y + a_2 x^2 + b_2 y^2\right)$
#. $\frac{\partial}{\partial x} \left( a_0 + a_1 x + b_1 y + c x y + a_2 x^2 + b_2 y^2\right)$
#. $\frac{\partial}{\partial x} A x^n y^m$
#. $\frac{\partial}{\partial y} A x^n y^m$
#. $\frac{\partial}{\partial x} \left( f(x) + y \right)$
#. $\frac{d}{dx} \left( f(x) + y \right)$
#. $\frac{d}{dx} \left( f(x+y)  \right)$
#. $\frac{\partial}{\partial x} \left( f(x) + g(y) \right)$
#. $\frac{\partial}{\partial y} \left( f(x) + g(y) \right)$
#. $\frac{d}{dx} \left( f(x) + g(y) \right)$
#. $\frac{\partial}{\partial x} f(x) g(y)$
#. $\frac{\partial}{\partial y} f(x) g(y)$
#. $\frac{d}{dx} f(x) g(y)$
#. $\frac{\partial}{\partial x} h(x,y) g(y)$
#. $\frac{\partial}{\partial y} h(x,y) g(y)$
#. $\frac{\partial}{\partial x} f(x) h(x,y) g(y)$
#. $\frac{\partial}{\partial y} f(x) h(x,y) g(y)$

# Exercise 39: Derivatives of Modeling Functions

Find each of these derivatives:

1. $\frac{d}{dt} \cos( \frac{2 \pi}{P} t )$
#. $\frac{d}{dt} \cos( a t )$
#. $\frac{d}{dt} \cos( t / T )$
#. $\frac{d}{dx} \ln(1+x)$
#. $\frac{d}{dx} \ln(1+a x)$
#. $\frac{d}{dx} \ln(b+a x)$
#. $\frac{d}{dt} e^{kt}$
#. $\frac{d}{dt} e^{t/T}$
#. $\frac{d}{dt} e^{kt} \cos(\frac{2 \pi}{P} t)$
#. Watch out! $\frac{d}{dx} \cos( a t )$  
#. $\frac{d}{dt} 2^{t/T}$
#. $\frac{d}{dt} (1+r)^{t}$
#. Watch out! $\frac{d}{dr} (1+r)^{t}$


# Exercise 40: Commit to memory

Here is a table of the derivatives and anti-derivatives of the basic modeling functions.  To make the forms as simple as possible, the functions have been written without parameters.  Of course, in actual practice, to use these functions for modeling you would need to write them with their appropriate parameters.  This table merely highlights one aspect of the patterns of derivatives.


  | Anti-Derivative | Function | Derivative
:---|:----:|:----:|:----: 
Name convention | F | f | f'
Operator notation  | $\int f(x) dx$ | $f(x)$ | $\frac{df}{dx}$
Basic Modeling | Functions  | 
Power Law | $\frac{1}{n+1} x^{n+1}$ | $x^n$ | $n x^{n-1}$
Sine | $-\cos(x)$ | $\sin(x)$ | $\cos(x)$
Cosine | $\sin(x)$ | $\cos(x)$ | $-\sin(x)$
Log | $x \ln(x) - x$ | $\ln(x)$ | $1/x$
Exponential | $e^x$ | $e^x$ | $e^x$


By memorizing the table, you will be able to reserve your intellectual capabilities for problems that matter, rather than struggling to derive routine facts.

1. After you've memorized the table above, fill in this version without referring to the original

  | Anti-Derivative | Function | Derivative
:---|:----:|:----:|:----:
Name convention | |  | 
Operator notation  | |
Basic Modeling | Functions
Power Law |  | |
Sine |  | | 
Cosine |  | | 
Log |  | |
Exponential |  | | 

#. Do it again!  But don't turn the sheet over to look.

  | Anti-Derivative | Function | Derivative
:---|:----:|:----:|:----:
Name convention | |  | 
Operator notation  | |
Basic Modeling Functions |
Power Law |  | |
Sine |  | | 
Cosine |  | | 
Log |  | |
Exponential |  | | 

#. Using the relationships shown in the "Function" and "Derivative" columns, show that differentiating the expression in the Anti-Derivative column leads to the expression in the Function column.  To differentiate the Anti-Derivative of the log function, you will need to use the product rule: $f(x) g(x) \rightarrow f'(x) g(x) + f(x) g'(x)$.

# Exercise 41

One of the important techniques for computing a derivative involves symbolic manipulation.  This is not usually a process that involves deep thought, but rather one that calls for good pattern recognition skills and the systematic application of a small set of rules.  There are some symbolic arithmetic problems that everybody should be able to do fluently and others that, while doable, are so involved that there is a substantial risk of making a mistake.  The situation is quite similar to that of arithmetic.  All numerically literate people should be able easily to solve problems like $3 + 9$ or $15/3$ or $7\times 2$.  Most people, even quite literate ones, cannot easily do similar problems such as $3592.35 + 5823.28$ or $15.32 / 2.87$.

"Basic problems" are ones you should be able to do fluently without much thought.  Some examples:


1. $\frac{d}{dx} x$

    a. $0$
    #. *$1$*
    #. $x$
    #. $x^2$


#. $\frac{d}{dx} k$

    a. *$0$*
    #. $1$
    #. $x$
    #. $x^2$


#. $\frac{d}{dx} k$

   a. *$0$*
   #. $1$
   #. $k$
   #. $kx$


#. $\frac{d}{dx} k x$

    a. $0$
    #. $1$
    #. *$k$*
    #. $kx$


#. $\frac{d}{dx} k x^2$

     a. $0$
     #. $k$
     #. $2 k$
     #. $kx$
     #. *$2 k x$*


#. $\frac{\partial}{\partial t} k x^2$ (Note the differentiation with respect to $t$.)

     a. *$0$*
     #. $k$
     #. $2 k$
     #. $kx$
     #. $2 k x$


#. $\frac{\partial}{\partial t} k x^2 t$ (Note the differentiation with respect to $t$.)

    a. $0$
    #. $k$
    #. $2 k x$
    #. *$k x^2$*
    #. $2 k x^2$


#. $\frac{\partial}{\partial t} e^t$ 

    a. $0$
    #. $t$
    #. *$e^t$*
    #. $t e^t$


#. $\frac{\partial}{\partial t} k e^t$ 

    a. $0$
    #. $t$
    #. *$k e^t$*
    #. $k e^{k t}$
    #. $t e^k$


#. $\frac{\partial}{\partial t} e^{kt}$ 

   a. $0$
   #. $t$
   #. $k e^t$
   #. *$k e^{k t}$*
   #. $t e^k$


#. $\frac{\partial}{\partial t} (1+r)^t$ 

    a. $(1+r)^t$
    #. *$\log(1+r) \cdot (1+r)^t$*
    #. $(1+r) \cdot (1+r)^t$
    #. $(1+t) \cdot (1+r)^t$


#. $\frac{\partial}{\partial r} (1+r)^t$ (Note that $r$ is the variable of differentiation.)

     a. $(1+r)^t$
     #. *$(1+r)^{t-1}$*
     #. $(1+t) \cdot (1+r)^t$



#. $\frac{\partial}{\partial t} \sin( \frac{2\pi}{P} t )$ 

     a. $\cos( \frac{2 \pi}{P} t$
     #. *$\frac{2 \pi}{P} \cos( \frac{2 \pi}{P} t$*
     #. $\sin( \frac{2 \pi}{P} t$
     #. $\frac{2 \pi}{P} \sin( \frac{2 \pi}{P} t$


#. $\frac{d}{dx} f(a x + b)$  where $\frac{d}{dx} f(x) = f'(x)$.

    a. $f'(a x + b)$
    #. *$a f'(a x + b)$*
    #. $b f'(a x + b)$
    #. $(a+b) f'(a x + b)$


#. $\frac{\partial}{\partial x} a_0 + a_1 x + a_2 y + a_3 x y + a_4 x^2 + a_5 y^2$

    a. $a_1 + 2 a_4 x$
    #. *$a_1 + a_3 y + 2 a_4 x$*
    #. $a_1 + a_3 x + 2 a_4 x$
    #. $a_1 + a_3 x + 2 a_4 x + a_5 y$


#. $\frac{\partial^2}{\partial x^2} a_0 + a_1 x + a_2 y + a_3 x y + a_4 x^2 + a_5 y^2$

    a. $0$
    #. $a_4$
    #. *$2 a_4$*
    #. $2 a_4 + 2 a_5$


#. $\frac{\partial^2}{\partial  y \partial x} a_0 + a_1 x + a_2 y + a_3 x y + a_4 x^2 + a_5 y^2$

    a. $0$
    #. $a_4$
    #. *$a_3$*
    #. $2 a_4 + 2 a_5$



