---
title: "Exercises on Solving"
author: "Daniel Kaplan"
---

```{r include=FALSE}
library(mosaic)
library(mosaicCalc)
library(directlabels)
```

# Exercise 1
 
By examining a plot of the appropriate function, find a solution to each of these equations. Zoom in on the plot to locate the solution to 3 decimal places.

1. $4 + 2 x = 3$
2. $x^2 - 7 x + 5 = 0$
3. $2 x^2 + 5 = 4 x$
4. $1 = \frac{1}{x} +\frac{1}{x^2} + \frac{1}{x^3} + \frac{1}{x^4} \ \mbox{for}\ x \in [1, 5]$.
5. $| \sin x^2 | = 2 e^{-2 x^2} \ \mbox{for} \ x \in [-5, 5]$.
6. $\sin \sqrt{5 + 7 x} = \cos x^2 \ \mbox{for} \ x \in [1, 5]$.

# Exercise 2

By examining a plot of the appropriate function, find a solu- tion to each of these equations. Use `mosaicCalc::findZeros()` to identify the solution to five decimal places.

1. $4 + 2 x = 3$ 
2. $x^2 - 7 x + 5 = 0$
3. $2 x^2 + 5 = 4 x$
4. $1 = \frac{1}{x} +\frac{1}{x^2} + \frac{1}{x^3} + \frac{1}{x^4} \ \mbox{for}\ x \in [1, 5]$.
5. $| \sin x^2 | = 2 e^{-2 x^2} \ \mbox{for} \ x \in [-5, 5]$.

# Exercise 3

Finding the zeros of a function of two variables is easy graphically: make a contour plot and find the zero contour! Any point on the zero contour is a solution.

For each of the following equations involving two variables, find a solution by constructing an appropriate function and finding a zero of it. Plug your solution into the function to demonstrate that it is correct:

1. $3 x + y = 7$ for $x \in [-5, 5]$ and $y \in [-5, 5]$.
#. $4 x^2 = 5 - y$ for $x \in [-5, 5]$ and $y \in [-5, 5]$.
#. $\sin x = \tan y^3$ for $x \in [-5, 5]$ and $y \in [-1, 1]$.


# Exercise 4

"Solving simultaneous equations" is the process of finding solutions that make more than one equation true at the same time. In general, a solution to single equation in two variables will be a point on a contour line. Any other point on that contour line will also be a solution, so there is not a unique solution. (That's not problematic. It just means that there’s more than one way to skin the proverbial cat.)

If there are two equations that need to be satisfied simultaneously, you can start by finding a solution to one of the equations and tracing out that contour line. Then do the same for the other equation. The simultaneous solution will be a point that is on both of the solution contour lines: the place where the contours cross.

Hint: Plot out the first function in the normal way. When adding on the second function to the plot, you will find it easier to read if you do not fill the space between the contour lines, and if you color the contour lines, say, red. You may also want to simplify the graphic by drawing only the zero contour. The argument `breaks = ` does this. For instance:

```{r results="hide", message = FALSE}
library(mosaic)
library(mosaicCalc)
graphFun(sin(x) - tan(y^3) ~ x + y,
         xlim = range(-5, 5), ylim = range(-1, 1),
         breaks = 0) %>%
  graphFun(x^2 + y^2 - 3 ~ x + y,
           xlim = range(-5, 5), ylim = range(-1, 1),
           tile = FALSE, color = "red", breaks = 0) 
```

Find a simultaneous solution to each of the following pairs of equations:

1. $3 x + y = 7$ and $x^2 + y^2 = 3$ for $x \in [-5, 5]$ and $y \in [-5, 5]$.
2. $4 x^2 = 5 - y$ and  $2 x^2 - y^2 = 4$ for $x \in [-5, 5]$ and $y \in [-5, 5]$.
3. $\sin x = \tan y^3$ and $x^2 + y^2 = 3$ for $x \in [-5, 5]$ and $y \in [-1, 1]$.

