---
chapter: "Functions"
author: "Danny Kaplan"
output: learnr::tutorial
tutorial:
  id: "spruce-think-plate"
  version: 0.1 DRAFT
runtime: shiny_prerendered
date: 2020-05-07 
tags: [R syntax, math notation, makeFun, sine, exp]
id: "spruce-think-plate" 
title: "Oscillations as they die out"
---

(ref:spruce-think-plate) Exercise in file spruce-think-plate

```{r spruce-think-plate-default, include = FALSE}
SDSdata::sds_setup()
library(learnr)
library(mosaicCalc)
library(gradethis)
tutorial_options(exercise.timelimit = 60, exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
```

TITLE GOES HERE: Here is a function of time, $t$, that gives a simple model of the vibrational response of an object when struck. (For instance, the plucking of a guitar string or the oscillations in a car's suspension after going over a bump.)

$$\sin( 2 \pi t / P ) e^{-k t}$$

Your job is to use `makeFun()` to create an R function that corresponds to the above expression. Give the function the name `wave`.

Some things to think about:

1. Multiplication must be explicitly written in R statements. The multiplication implicit in mathematical notation expressions like $2 \pi t$ does not apply in R.
2. In R (and most other computer languages), the exponential function is called `exp()`.
3. We said that the expression is a function of time $t$. Put there are also *parameters* and *constants* in the expression. When evaluating the function or making a graph, numerical values must be assigned to each parameter.


```{r spruce-think-plate-box-setup}

```

```{r spruce-think-plate-box, exercise=TRUE, exercise.setup="spruce-think-plate-box-setup", exercise.cap="Code box", exercise.eval=FALSE, exercise.lines=2, exercise.completion=FALSE, exercise.startover=TRUE, exercise.diagnostics=FALSE }
# Define your "wave" function here.
waave <- function(t) {t^2}

```

```{r spruce-think-plate-box-check, echo = FALSE}
# Checking code goes here.
TRUE
```

```{r spruce-think-plate-box-code-check-inactive, echo = FALSE}
# Code checking prior to execution of code

```


```{r spruce-think-plate-box-solution, echo = FALSE}
wave <- sin(2 * pi * t / P) * exp(- k * t)
```

1. Graph the `vibration()` function over a domain from t=0 to t = 10. Set the parameter $P$ to 1.5 and the parameter $k$ to 0.1. 

MULTIPLE CHOICE QUESTION HERE.

2. Leaving $k$ at 0.1, make a contour plot of `vibration()` as a function of $t$ and $P$. The domain for $P$ in the graphic should be 0.1 to 2. 

MULTIPLE CHOICE QUESTION about contour plot.
```{r spruce-think-plate-Q1, echo=FALSE}
question(
  text = "The prompt for the question",
  answer("First choice", correct = TRUE, message = "The first one is *always* correct."),
  answer("Second choice", message = "The second choice is *never* correct."),
  answer("Third choice", correct = TRUE, message = "More than one answer can be true."),
  allow_retry = TRUE,
  random_answer_order = TRUE)
```

