---
title: "Math 141Z Assignment 1"
output: learnr::tutorial
tutorial:
  id: "Assignment-1"
  version: 0.1
runtime: shiny_prerendered
date: 2020-06-17
tags: [first, second, third]
id: Assignment-1
---

```{r ant-feel-magnet-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(submitr)
library(basket)
library(USAFAexercises)


learnr::tutorial_options(exercise.timelimit = 60,
                 exercise.checker = gradethis::grade_learnr)
```

```{r echo=FALSE}
submitr::login_controls()
```

```{r context="server", echo = FALSE}
options(tutorial.storage = "none")
vfun <- basket::check_valid
  #submitr::make_basic_validator(NULL, "hello")
storage_actions <- submitr::record_gs4("1w3fEld2rZlR_6FuzvkA-viOLBA5JdqD3xHl-LuLX3-Y", "statprep.annie@gmail.com", vfun)
  # submitr::record_local("./minimal_submissions.csv")
submitr::shiny_logic(input, output, session, vfun,
                     storage_actions)
```


<!--  begin contents here -->

`r  prob_count <- 0`

## Problem `r (prob_count <- prob_count + 1)`

`r etude::include_etude(system.file("Exercises/Functions/deer-rise-mug.Rmd", package = "USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`r etude::include_etude(system.file("Exercises/Functions/shark-leave-canoe.Rmd", package = "USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/horse-cost-blanket.Rmd", package = "USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/chicken-choose-vase.Rmd",  package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/wolf-lead-kitchen.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/eagle-speak-pan.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/giraffe-catch-pot.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/crow-mean-dress.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/cat-lend-futon.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/lamb-talk-gloves.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/snake-choose-kayak.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/cow-hang-closet.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/snail-hold-drawer.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/finch-feel-knob.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/sheep-drink-vase.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Modeling_functions/wolf-hurt-blanket.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Modeling_functions/goldfish-draw-plant.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/giraffe-understand-pants.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Integrals/beech-sleep-drawer.Rmd", package="USAFAexercises"))`

## Problem `r (prob_count <- prob_count + 1)`

`fr etude::include_etude(system.file("Exercises/Functions/ash-send-shirt.Rmd", package="USAFAexercises"))`
