---
chapter: "none yet"
title: "Title for goat-pitch-socks"
author: "Daniel Kaplan"
difficulty: "XX"
output: learnr::tutorial
tutorial:
  id: "goat-pitch-socks"
  version: 0.1
runtime: shiny_prerendered
date: 2020-12-02 
tags: [deriv-anti-deriv, second, third]
id: goat-pitch-socks
---

```{r goat-pitch-socks-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` Choose the correct term for ____ in each of the following:


```{r goat-pitch-socks-QA1, echo=FALSE}
etude::choose_one(
  prompt = "Velocity is the _____ of acceleration (w.r.t. time)?",
  choices = list(
    "2nd deriv" = "",
    "deriv" = "",
    "+anti-deriv+" = "",
    "2nd anti-deriv" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```

```{r goat-pitch-socks-QA2, echo=FALSE}
etude::choose_one(
  prompt = " Position is the _____ of velocity (w.r.t. time)?",
  choices = list(
    "2nd deriv" = "",
    "+deriv+" = "",
    "anti-deriv" = "",
    "2nd anti-deriv" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```



```{r goat-pitch-socks-QA3, echo=FALSE}
etude::choose_one(
  prompt = "Acceleration is the _____ of position (w.r.t. time)?",
  choices = list(
    "+2nd deriv+" = "",
    "deriv" = "",
    "anti-deriv" = "",
    "2nd anti-deriv" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```



```{r goat-pitch-socks-QA4, echo=FALSE}
etude::choose_one(
  prompt = "Energy is the ______ of force (w.r.t. time): ",
  choices = list(
    "2nd deriv" = "",
    "deriv" = "",
    "anti-deriv" = "",
    "2nd anti-deriv" = "",
    "+none of the above+" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```


```{r goat-pitch-socks-QA5, echo=FALSE}
etude::choose_one(
  prompt = " Energy is the _____ of force (w.r.t. distance)?",
  choices = list(
    "2nd deriv" = "",
    "deriv" = "",
    "+anti-deriv+" = "",
    "2nd anti-deriv" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```


```{r goat-pitch-socks-QA6, echo=FALSE}
etude::choose_one(
  prompt = "Cash is the _____ of cash flow (w.r.t. time)?",
  choices = list(
    "2nd deriv" = "",
    "deriv" = "",
    "+anti-deriv+" = "",
    "2nd anti-deriv" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```


```{r goat-pitch-socks-QA7, echo=FALSE}
etude::choose_one(
  prompt = "Acceleration is the _____ of velocity (w.r.t. time)?",
  choices = list(
    "2nd deriv" = "",
    "+deriv+" = "",
    "anti-deriv" = "",
    "2nd anti-deriv" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```

```{r goat-pitch-socks-QA8, echo=FALSE}
etude::choose_one(
  prompt = " Position is the _____ of velocity (w.r.t. time)?",
  choices = list(
    "2nd deriv" = "",
    "deriv" = "",
    "+anti-deriv+" = "",
    "2nd anti-deriv" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```



```{r goat-pitch-socks-QA9, echo=FALSE}
etude::choose_one(
  prompt = "Distance is the _____ of acceleration (w.r.t. time)?",
  choices = list(
    "2nd deriv" = "",
    "deriv" = "",
    "anti-deriv" = "",
    "+2nd anti-deriv+" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```



```{r goat-pitch-socks-QA10, echo=FALSE}
etude::choose_one(
  prompt = "Force is the ______ of momentum (w.r.t. time): ",
  choices = list(
    "2nd deriv" = "",
    "+deriv+" = "",
    "anti-deriv" = "",
    "2nd anti-deriv" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```


```{r goat-pitch-socks-QA11, echo=FALSE}
etude::choose_one(
  prompt = "Force is the _____ of energy (w.r.t. distance)?",
  choices = list(
    "2nd deriv" = "",
    "+deriv+" = "",
    "anti-deriv" = "",
    "2nd anti-deriv" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```


```{r goat-pitch-socks-QA12, echo=FALSE}
etude::choose_one(
  prompt = "Cash is the _____ of cash flow (w.r.t. time)?",
  choices = list(
    "2nd deriv" = "",
    "deriv" = "",
    "+anti-deriv+" = "",
    "2nd anti-deriv" = "",
    "none of the above" = ""
  ),
  random_answer_order = FALSE,
  inline = FALSE
)
```

Circumference of a circle is the derivative of area with respect to radius.

1. Energy is the ............. of force with respect to distance. {antideriv}{deriv, 2nd deriv, antideriv, none of these}

2. Power is the ............. of energy with respect to time. {deriv}{deriv, 2nd deriv, antideriv, none of these}

3. Total cost is the ............. of marginal cost with respect to quantity. {antideriv}{deriv, 2nd deriv, antideriv, none of these}

<!--begin-answer-->  
<!--Really this should be "for the homework version"-->
4. Births per year is part of the ............. of population size with respect to time. (The other parts relate to deaths, immigration, and emigration.)
{deriv}{deriv, 2nd deriv, antideriv, none of these}
<!--end-answer-->


1. Position is the ............. of velocity with respect to time.
{antiderivative}{deriv, 2nd deriv, antideriv, none of these}

2. Acceleration is the ............. of velocity with respect to time. {deriv}{deriv, 2nd deriv, antideriv, none of these}

3. Acceleration is the ............. of position with respect to time. {2nd deriv}{deriv, 2nd deriv, antideriv, none of these}

<!--begin-answer-->} 
<!-- Really this should be "for the homework version"-->
4. Density is the ............. of mass with respect to position. {deriv}{deriv, 2nd deriv, antideriv, none of these}
<!--end-answer-->
