---
chapter: "none yet"
title: "Title for ash-send-shirt"
author: "Danny Kaplan"
output: learnr::tutorial
tutorial:
  id: "ash-send-shirt"
  version: 0.1
runtime: shiny_prerendered
date: 2020-06-05 
tags: [first, second, third]
id: ash-send-shirt
---

```{r ash-send-shirt-setup, include = FALSE}
SDSdata::sds_setup()
library(etude)
library(learnr)
library(gradethis)

library(ggformula)
library(mosaic)
library(mosaicCalc)
library(mosaicUSAFA)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = QandA::save_check_inputs)
```


`r etude::exercise_title()` Finding the zeros of a function of two variables is easy graphically: make a contour plot and find the zero contour! Any point on the zero contour is a solution.

For each of the following equations involving two variables, find a solution by constructing an appropriate function, making a contour plot, and locating a point on the zero contour. Plug your solution into the function to demonstrate that it is correct. 


```{r ash-send-shirt-C1, exercise=TRUE, exercise.cap="Sandbox"}

```

a. $3 x + y = 7$ for $x \in [-5, 5]$ and $y \in [-5, 5]$.
b. $4 x^2 = 5 - y$ for $x \in [-5, 5]$ and $y \in [-5, 5]$.
c. $\sin x = \tan y^3$ for $x \in [-5, 5]$ and $y \in [-1, 1]$.


Of course, reading coordinates off a graph is not necessarily precise, so even if your coordinates are reasonable, the function value at those coordinates may not be exactly zero. But make sure that it is small.

```{r ash-send-shirt-C3}
etude::essay_response(
  "Enter the coordinate you found for each of functions (a), (b), and (c)."
)
```

```{r ash-send-shirt-C2}
etude::choose_one(
  "In deciding whether coordinates you read off the graph for a function zero are reasonable, to what should you compare the function output at those coordinates?",
  list(
    "to zero" = "Obvious though this answer is, it's not useful. The same non-zero value can be either very close to zero or very far from zero depending on the context. For example, the distance from the Earth to the Sun is about $1.5 \\times 10^{-5}$ light years. That's a small number, but you wouldn't be alive if the number were close to zero!",
    "to one" = "One nanometer is a distance that's small on a human scale but large on an atomic scale.",
    "+small compared to the value of the nearest non-zero contour+" = "",
    "no way to know" = "You can know. But the answer is not a matter of arithmetic but of knowing what's important in the appropriate context of your problem."
  ),
  random_answer_order = FALSE
)
```

