---
chapter: "none yet"
title: "Title for crow-mean-dress"
author: "Danny Kaplan"
output: learnr::tutorial
tutorial:
  id: "crow-mean-dress"
  version: 0.1
runtime: shiny_prerendered
date: 2020-06-03 
tags: [power-law, allometrics, third]
id: crow-mean-dress
---

```{r crow-mean-dress-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(ggformula)
library(mosaic)
library(mosaicCalc)
library(math141Z)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` The surface area $S$ of a mammal is reasonably well approximated by the function $S(M) \equiv k M^{2/3}$ where $M$ is the body mass and the constant $k$ depends on the particular species under consideration. 


```{r crow-mean-dress-Qinline1, echo=FALSE}
etude::choose_one(
  prompt = "A) Consider a baby and an adult. The adult's mass is $8$ times greater than the baby's. Then the adult's surface area is ...?",
  c("The same as the baby's", "1.5 times of the baby's", "+4 times the baby's+", "8 times the baby's"),
  random_answer_order = FALSE,
  inline = FALSE
)
```


<span title='crow-mean-dress-sandbox1'>...</span>
```{r crow-mean-dress-sandbox1, exercise=TRUE, exercise.cap="Sandbox", exercise.setup="crow-mean-dress-sandbox1-setup", exercise.eval=FALSE, exercise.lines=2, exercise.completion=FALSE, exercise.startover=TRUE, exercise.diagnostics=TRUE}

```

```{r crow-mean-dress-QA1, echo=FALSE}
# indicate correct choices with +_+ in the name of the list item.
etude::choose_one(
  prompt = "B) Consider a human of body mass 70 kg with a skin surface area of 18,600 cm^2^. Which of the following units for the constant of proportionality $k$ is correct?",
  choices = list(
    "+cm$^2$ kg$^{-2/3}$+" = "",
    "cm$^2$" = "What about the units of mass? What takes care of them?",
    "cm$^2$ kg$^{2/3}$" = "When you multiply 70 kg by $k$, you need to get a result in $cm^2$.",
    "kg$^{-1}$" = "Would this cancel out the units of mass and produce $cm^2$ for the result?"
  )
)
```

```{r crow-mean-dress-Qinline2, echo=FALSE}
etude::choose_one(
  prompt = "C)  In the units of part (B), which value is $k$ closest to?",
  c(1, 10, 100, "+1000+"),
  random_answer_order = FALSE
)
```

The numerical value of the constant $k$ changes depending on what units you want to express it in. The value you found in part (C) works for masses stated in kg and skin areas in cm$^2$. Suppose you want to present $k$ in units that will make sense to people who are used to talking about skin area in square inches and mass in pounds. 1 inch \approx 2.6 cm$.  In the rough-and-ready way everyday people express themselves, 1 kg \approx 2.2 lbs. (Of course, pounds is a measure of force, not mass. But people use it as if it were mass. A mass of 70 kg corresponds to about 4.8 slugs. In Earth's gravity, 4.8 slugs produces a force of 154 pounds.)

```{r crow-mean-dress-QA2, echo=FALSE}
etude::choose_one(
  prompt = "Optional challenge) What is the rough numerical value of the conversion factor which would produce a constant of proportionality that would be suitable for converting body mass in pounds to skin area in square-inches?",
  choices = c(0.03, "+0.3+", 3, 30, 300),
  random_answer_order = FALSE
)
```

```{etude}
$k$ is about 1000 cm$^2$ kg$^{-2/3}$. To convert the cm to inches, we need to multiply by (1 in/2.6 cm)$^2$. To convert the kg to pounds, we need to multiply by (2.2 pounds/ 1 kg)$^{-2/3}$.
```




