---
chapter: "none yet"
author: "Danny Kaplan"
difficulty: "XX"
date: 2020-07-19 
version: 0.1
tags: [wind chill, graphics, by-hand]
id: giraffe-understand-bulb
---

```{r giraffe-understand-bulb-default, include = FALSE}
library(etude)
library(mosaicCalc)
library(ggformula)
```

`r etude::exercise_title()`  "Wind chill" is a quantity used to describe how cold a person feels.  It takes into account both ambient air temperature and the wind speed. The official definition is based on empirical engineering relationships of heat transfer in a moving fluid as defined in 2001 by the US/Canada/UK Joint Action Group for Temperature Indices (JAG/TI). For convenience, it has been approximated by a formula

$$T_{wc} \equiv 13.12 + 0.6214 T_a - 11.37 v^{0.16} + 0.3965 T_a v^{0.16} ,$$
where $T_a$ is ambient temperature in deg. C and $v$ is wind velocity measured in kilometers per hour. (The measurement should be made at 10m, the standard anemometer height.)

The figure shows wind chill as a function of air temperature and wind speed.

```{r echo = FALSE}
wind_chill <- function(T_ambient, v_wind) {
  13.12 + 0.6215 * T_ambient - 
    11.37 * v_wind^0.16 + 
    0.3965 * T_ambient * v_wind^0.16
}
contour_plot(wind_chill(T_ambient, v_wind) ~ T_ambient + v_wind,
             domain(T_ambient = c(-50, 0), v_wind = c(10, 50)),
             skip = 0, filled = FALSE, n_contours = 20, 
             contour_color = "gray", label_color = "black") %>%
  gf_labs(title = "Wind Chill (deg C)", y = "Wind velocity (km/hr)",
          x = "Ambient temperature (deg C)") %>%
  gf_theme(theme_bw())

```

The following questions involve reading quantities off the graph. In addition to giving the numerical answer called for, you should mark a point on the graph that is relevant to the answer you are giving. For each item, use the letter label for that item, e.g., (b), as your mark. 

a. What is the wind chill if the air temperature is $-25^{\circ}$C and the wind speed is 15 kph?  Mark the relevant point on the graph with label "(a)" and give the numerical value (with units) of wind chill.


#. Bare skin is at risk of frostbite in just a few minutes when the wind chill is below $-45^{\circ}$C.  If the air temperature is $-30^\circ$C, what is the wind speed at which the wind chill is $-45^{\circ}$? Mark the point on the relevant point on the graph with label "(b)" and give the numerical value of wind speed.


#. Find a point where the wind speed produces a wind chill 20 degrees colder than the air temperature?  Mark it clearly on the graph with the label  "(c)".  





