---
chapter: "none yet"
title: "Title for lamb-talk-gloves"
author: "Danny Kaplan"
output: learnr::tutorial
tutorial:
  id: "lamb-talk-gloves"
  version: 0.1
runtime: shiny_prerendered
date: 2020-06-03 
tags: [first, second, third]
id: lamb-talk-gloves
---

```{r lamb-talk-gloves-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(ggformula)
library(mosaic)
library(mosaicCalc)
library(math141Z)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```

`r etude::exercise_title()` A person breathes in and out every three seconds. The volume $V$ of air in the person's lungs varies between a minimum of $2$ liters and a maximum of $4$ liters. Assume time $t$ is measured in seconds.

Remember that a full cycle of the sine wave $\sin(x)$ involves $x$ going  from its starting value to that value **plus** $2 \pi$.  

```{r lamb-talk-gloves-QA1, echo=FALSE}
wrong_amp <- "This varies between a minimum of 0 and a maximum  of 2."
etude::choose_one(
  prompt = "Which of the following is the most appropriate of these models for $V(t)$?",
  choices = list(
    "$V(t) \\equiv 2 + 2 \\sin \\left( \\frac{\\pi}{3} t \\right)$" = wrong_amp,
"+$V(t) \\equiv 3 + \\sin \\left( \\frac{2\\pi}{3} t \\right)$+" = "Good. In this class, we generally write the sine function like  $\\sin(2 \\pi t/P)$ which means that the overall argument to  the sine function will go from 0 to $2 \\pi$ when $t$ goes from 0 to $P$.",
"$V(t) \\equiv 2 + 2 \\sin \\left( \\frac{2\\pi}{3} t \\right)$" = wrong_amp,
"$V(t) \\equiv 3 + \\sin \\left( \\frac{\\pi}{3} t \\right)$" = "Right amplitude and baseline: the minimum will be 2 liters and the maximum 4 liters. But the period is wrong. Going from $t=0$ to  $t=3$ should produce a full cycle of the sine function. But here  the argument would go only  from 0 to $\frac{\\pi}{3} 3 = $\\pi$. Thus,  after 3 seconds, only half  a cycle has been  completed."
  ),
  inline = FALSE,
  random_answer_order = FALSE
)
```


```{r lamb-talk-gloves-QA2, echo=FALSE}

etude::choose_one(
  prompt = "Given a respiratory period of 3 seconds/breath, what is the respiratory **frequency** in units of breaths/minute?",
  choices = list(
    "+20 breaths/minute+" = "Right. Each breath takes 1/20th of a minute, which is 3 seconds, the period specified in the question.",
    "3 breaths/minute" = "If this were true, each breath would take 20 seconds to complete.",
    "1/3 breath per minute" = "With breaths completed every three seconds, 1/3 of a breath is completed each second. But the problem asked for breaths per minute.",
    "20 seconds per breath" = "The *period* is in the units of seconds per breath, but the *frequency* will have units of breaths per second. Frequency is the reciprocal of period (and vice versa). "
  ),
  inline = FALSE
)
```

A respiratory cycle can be divided into two parts: inspiration and expiration. Please do an experiment. Using a clock or watch, breath with a total period of 3 seconds/breath, that is, complete one breath every three seconds. Once you  have practiced this and can do it without forcing either phase of breathing,  make a rough estimate of what fraction of the cycle is inspiration and what fraction is expiration. (The "correct/incorrect" answers here are right for most people. Your natural respiration might be different.)

```{r lamb-talk-gloves-QA3, echo=FALSE}
etude::choose_one(
  prompt = "Which is true?",
  choices = list(
    "+Inspiration lasts longer than expiration+" = "",
    "Expiration lasts longer than inspiration" = "Maybe it is for you, but not for most people. Try breathing in while counting 1-2-3 then exhaling while counting 1-2-3-4-5-6. Likely,  that's not a very  natural  pattern for you.",
    "Inspiration and expiration each consume about  the same fraction of the complete  cycle." = "People can do this consciously by counting 1-2-3  for inspiration  and another 1-2-3 for expiration. This usually feels forced and unnatural."
  ),
  inline = FALSE
)
```



