---
chapter: "Functions"
title: "Title for snake-choose-kayak"
author: "Danny Kaplan"
output: learnr::tutorial
tutorial:
  id: "snake-choose-kayak"
  version: 0.1
runtime: shiny_prerendered
date: 2020-06-04 
tags: [two-variables, economics, third]
id: snake-choose-kayak
---

```{r snake-choose-kayak-setup, include = FALSE}
library(etude)
library(learnr)
library(gradethis)
library(ggformula)
library(mosaic)
library(mosaicCalc)
library(mosaicUSAFA)

knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(exercise.timelimit = 60, 
                 exercise.checker = gradethis::grade_learnr)
```


`r etude::exercise_title()` In economic theory, the quantity of the demand for any good is a decreasing function of the price of that good and an *increasing* function of the price of a competing good.

The classical example is that apple juice competes with orange juice. The quantity of demand for orange juice might be measured in, say, thousands of liters of orange juice.

```{r}
f <- function(x,y) {
  angle = atan2(y, x)
  300* ((pi/2) - angle)
}
contour_plot(f(x, y) ~ x + y, domain(x=c(3,1.3), y=c(.4,1.3)),
             skip = 0) %>%
  gf_labs(title = "Demand quantity", x = "?", y = "?")
```

Note that the axes have not been labeled. You'll have to figure things out yourself.

```{r snake-choose-kayak-QA1, echo=FALSE}
explain1 <- "The prices of apple juice and of orange juice are the *inputs* to the function.  Input values are displayed as position on an axis. The contour labels identify the output value for any input that corresponds to a point on the contour."
explain2 <- "A reasonable answer, but the function output is the quantity of *orange* juice."
etude::choose_one(
  prompt = "What are the numbers labeling the contours?",
  choices = list(
    "+Quantity of orange juice.+" = "",
    "Price of apple juice." = explain1,
    "Price of orange juice." = explain1,
    "Quantity of apple juice." = explain2
  )
)
```

```{r snake-choose-kayak-QA2, echo=FALSE}
etude::choose_one(
  prompt = "What does the y-axis measure?",
  choices = list(
    "+Price of orange juice+" = "",
    "Quantity of apple juice" = "B's a consonant, Bub.",
    "Quantity of orange juice" = "C is a consonant, Chub",
    "Price of apple juice" = "D is a consonant, Dear"
  )
)
```


```{etude}
There are two possibilities to consider:

1. x is the price of orange juice and y the price of apple juice.
2. the reverse

Let's start with possibility (1). Suppose the price of orange juice is $2. Follow the path from y = 0.4 to y = 1.2. Moving along that path, the demand for orange juice -- according to possibility (1) -- goes *down*. But a price increase for a competing good should, according to theory, lead to an *increase* in *demand* for the good itself. Possibility (1) is inconsistent with the theory. That leaves possibility (2), where y is the price of orange juice. Suppose the price is $0.80. Follow the horizontal path from x = 1.5 to x = 3.0. The demand for orange juice goes *up* along this path, as would be expected in economic theory with a rising price of a competing good (here, apple juice). 
``` 

.


